Distributional Trees and Forests

### 2017-06-24 pkg/disttree/inst/useR2017/slides.Rnw
plots: now colored title instead of colored lines in legend
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1490 255f4249-2263-4417-b1ea-8498779e27a7

  if(!(any(dt.par, df.par, g.par, all.par, all.var))){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "True parameters", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
  col = transpgrey, border = "transparent")
    #lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", lty = 1, col = 'grey')
    #lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", lty = 1, col = 'grey')
    legend('topleft', expression(mu %+-% sigma), bty = 'n')
  }

  # disttree
  if(dt.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.dt + sp$fitted.sigma.dt, rev(sp$fitted.mu.dt - sp$fitted.sigma.dt)),
  col = pallight["tree"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    #lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
           col = c('black','grey',pal["tree"], pallight["tree"]), lty = 1, bty = 'n')
  }

  # distforest
  if(df.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "distforest", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.df + sp$fitted.sigma.df, rev(sp$fitted.mu.df - sp$fitted.sigma.df)),
  col = pallight["forest"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = 'orange')
    #lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = 'orange')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
           col = c('black','grey',pal["forest"], pallight["forest"]), lty = 1, bty = 'n')
  }

  # gamlss
  if(g.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey",  main = "gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.g + sp$fitted.sigma.g, rev(sp$fitted.mu.g - sp$fitted.sigma.g)),
  col = pallight["gamlss"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = 'lightblue')
    #lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = 'lightblue')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')),  
           col = c('black','grey',pal["gamlss"], pallight["gamlss"]), lty = 1, bty = 'n')
  }

  # disttree vs. distforest vs. gamlss
  if(all.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree vs. distforest vs. gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2)
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\hat{\\mu}$ disttree'),
                        TeX('$\\hat{\\mu}$ distforest'), TeX('$\\hat{\\mu}$ gamlss')), 
           col = c('black',pal["tree"],pal["forest"],pal["gamlss"]), lty = 1, bty = 'n')
  }

  # variance: disttree vs. distforest vs. gamlss
  if(all.var){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree vs. distforest vs. gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'black')
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.dt + sp$fitted.sigma.dt, rev(sp$fitted.mu.dt - sp$fitted.sigma.dt)),
  col = pallight["tree"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = pal["tree"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = pal["tree"], lwd = 2)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.df + sp$fitted.sigma.df, rev(sp$fitted.mu.df - sp$fitted.sigma.df)),
  col = pallight["forest"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = pal["forest"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = pal["forest"], lwd = 2)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.g + sp$fitted.sigma.g, rev(sp$fitted.mu.g - sp$fitted.sigma.g)),
  col = pallight["gamlss"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = pal["gamlss"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = pal["gamlss"], lwd = 2)
    legend('topleft', c(TeX('$\\mu \\pm \\sigma$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ disttree'),
                        TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ distforest'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ gamlss')), 
           col = c('black',pallight["tree"],pallight["forest"],pallight["gamlss"]), lty = 1, bty = 'n')
  }
}

@@

%\bibliography{ref.bib}

@@

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.41cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<1->{
<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt='n', yaxt='n', ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<2->{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt='n', yaxt='n', ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\visible<3->{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@

@@

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.2cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@
\visible<1->{
<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt='n', yaxt='n', ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
@
}
\end{center}

\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
 \begin{center}
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@
\visible<2->{
<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt='n', yaxt='n', ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
@
}
\end{center}

@@

\setkeys{Gin}{width=0.74\linewidth}
<<plottree_dgp, fig=TRUE, echo=FALSE>>=
pal2 <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(fig, inner_panel = node_inner_ext,
     tp_args = list(FUN = identity, width = 18, fill = pal2[c(1, 3)]), 
     ip_args = list(fill = pal2[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1.4)
@
\end{center}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plottree_xyplot, fig=TRUE, echo=FALSE>>=
par(mar=c(5.1,4.1,4.1,3.1))
plot(y=odata$y, x=odata$x, ylab = "y", xlab = "x", col = "grey")
#plot(y=odata$y, x=odata$x, xaxt='n', yaxt='n', ann=FALSE, col = "slategray")
lines(x = odata$x, y = odata$mu, col = pal["forest"], lwd = 2.5, main = "")
polygon(c(odata$x, rev(odata$x)), c(odata$mu + odata$sigma, rev(odata$mu - odata$sigma)),
  col = pallight["forest"], border = "transparent")

#lines(x = odata$x, y = odata$mu + odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
#lines(x = odata$x, y = odata$mu - odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
@
\end{center}
\end{frame}

>>>>

  if(!(any(dt.par, df.par, g.par, all.par, all.var))){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "True parameters", cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
  col = transpgrey, border = "transparent")
    #lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", lty = 1, col = 'grey')
    #lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", lty = 1, col = 'grey')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$')), 
    #       col = c('black', transpgrey), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # disttree
  if(dt.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree", col.main = pal["tree"], cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.dt + sp$fitted.sigma.dt, rev(sp$fitted.mu.dt - sp$fitted.sigma.dt)),
  col = pallight["tree"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    #lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
    #                    TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
    #       col = c('black','grey', pal["tree"], pallight["tree"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # distforest
  if(df.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "distforest", col.main = pal["forest"], cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.df + sp$fitted.sigma.df, rev(sp$fitted.mu.df - sp$fitted.sigma.df)),
  col = pallight["forest"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = 'orange')
    #lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = 'orange')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
    #                    TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
    #       col = c('black','grey',pal["forest"], pallight["forest"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # gamlss
  if(g.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey",  main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.g + sp$fitted.sigma.g, rev(sp$fitted.mu.g - sp$fitted.sigma.g)),
  col = pallight["gamlss"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = 'lightblue')
    #lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = 'lightblue')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
    #                    TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')),  
    #       col = c('black','grey',pal["gamlss"], pallight["gamlss"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # disttree vs. distforest vs. gamlss
  if(all.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", xaxt="n", yaxt="n", xlab = "", ylab = "")
    title(main = "disttree", col.main = pal["tree"], cex.main = 1.2, font.main = 2, adj = 0.19)
    title(main = "distforest", col.main = pal["forest"], cex.main = 1.2, font.main = 2, adj = 0.5)
    title(main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, font.main = 2, adj = 0.8)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.35)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.65)

    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2)
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\hat{\\mu}$ disttree'),
    #                    TeX('$\\hat{\\mu}$ distforest'), TeX('$\\hat{\\mu}$ gamlss')), 
    #       col = c('black',pal["tree"],pal["forest"],pal["gamlss"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 13.95, expression(mu), bty = "n")
  }

  # variance: disttree vs. distforest vs. gamlss
  if(all.var){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", xaxt="n", yaxt="n", xlab = "", ylab = "")
    title(main = "disttree", col.main = pal["tree"], cex.main = 1.2, font.main = 2, adj = 0.19)
    title(main = "distforest", col.main = pal["forest"], cex.main = 1.2, font.main = 2, adj = 0.5)
    title(main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, font.main = 2, adj = 0.8)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.35)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.65)

    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'black')
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.dt + sp$fitted.sigma.dt, rev(sp$fitted.mu.dt - sp$fitted.sigma.dt)),
  col = pallight["tree"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = pal["tree"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = pal["tree"], lwd = 2)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.df + sp$fitted.sigma.df, rev(sp$fitted.mu.df - sp$fitted.sigma.df)),
  col = pallight["forest"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = pal["forest"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = pal["forest"], lwd = 2)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.g + sp$fitted.sigma.g, rev(sp$fitted.mu.g - sp$fitted.sigma.g)),
  col = pallight["gamlss"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = pal["gamlss"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = pal["gamlss"], lwd = 2)
    #legend('topleft', c(TeX('$\\mu \\pm \\sigma$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ disttree'),
    #                    TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ distforest'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ gamlss')), 
    #       col = c('black',pal["tree"],pal["forest"],pal["gamlss"]), lty = 1, bty = "n", lwd = 2.5)
  legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

}

@@

@@

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.41cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<1->{
<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<2->{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\visible<3->{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@

@@

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.2cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@
\visible<1->{
<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
@
}
\end{center}

\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
 \begin{center}
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@
\visible<2->{
<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
@
}
\end{center}

@@

\setkeys{Gin}{width=0.74\linewidth}
<<plottree_dgp, fig=TRUE, echo=FALSE>>=
pal2 <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(fig, inner_panel = node_inner_ext,
     tp_args = list(FUN = identity, width = 18, height = 5.002, fill = pal2[c(1, 3)]), 
     ip_args = list(fill = pal2[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1.4)
@
\end{center}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plottree_xyplot, fig=TRUE, echo=FALSE>>=
par(mar=c(5.1,4.1,4.1,3.1))
plot(y=odata$y, x=odata$x, ylab = "y", xlab = "x", col = "grey")
#plot(y=odata$y, x=odata$x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray")
lines(x = odata$x, y = odata$mu, col = pal["forest"], lwd = 2.5, main = "")
polygon(c(odata$x, rev(odata$x)), c(odata$mu + odata$sigma, rev(odata$mu - odata$sigma)),
  col = pallight["forest"], border = "transparent")
legend(x = -0.071, y = 21.6, expression(mu  %+-%  sigma), bty = "n")
#lines(x = odata$x, y = odata$mu + odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
#lines(x = odata$x, y = odata$mu - odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
@
\end{center}
\end{frame}
